---
title: "Stadard Errors"
author: "Corwin Bell"
date: "11/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(UsingR)
x = father.son$fheight
y = father.son$sheight
n = length(y)
N = 50
set.seed(1)
index = sample(n,N)
sampledat = father.son[index,]
x = sampledat$fheight
y = sampledat$sheight
betahat = lm(y~x)$coef
```

```{r}
fit = lm(y~x)
deviance(fit)
```
```{r}
y_hat = fit$fitted.values
SR = (y_hat-y)^2
SSR = sum(r_sq)
SSR
```

```{r}
sigma2 = SSR / 48
```
Form the design matrix X (note: use a capital X!). This can be done by combining a column of 1's with a column of 'x' the father's heights.
```{r}
X = cbind(rep(1,N), x)
X
```
Now calculate (ğ‘‹ğ‘‡ğ‘‹)âˆ’1, the inverse of X transpose times X. Use the solve() function for the inverse and t() for the transpose. What is the element in the first row, first column?
```{r}
X_inv = solve(t(X)%*%X)
X_inv
```
Now we are one step away from the standard error of beta-hat. Take the diagonals from the (ğ‘‹ğ‘‡ğ‘‹)âˆ’1 matrix above, using the diag() function. Now multiply our estimate of ğœ2 and the diagonals of this matrix. This is the estimated variance of beta-hat, so take the square root of this. You should end up with two numbers, the standard error for the intercept and the standard error for the slope.
What is the standard error for the slope?
```{r}
sqrt(sigma2 * diag(X_inv))
```
```{r}
n <- nrow(father.son)
N <- 50
index <- sample(n,N)
sampledat <- father.son[index,]
x <- sampledat$fheight
y <- sampledat$sheight
X <- model.matrix(~x)

N <- nrow(X)
p <- ncol(X)

XtXinv <- solve(crossprod(X))

resid <- y - X %*% XtXinv %*% crossprod(X,y)
ssr = sum(resid^2)
s <- sqrt( sum(resid^2)/(N-p))
ses <- sqrt(diag(XtXinv))*s
ssr
```

